Supplementary Information

### 2023-11-19 supplement.Rmd
ismail modifs psych science

Participants played a 15-round repeated trust game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. As seen in Figure \ref{fig:trustGameScreenshot} each player was represented with an icon with the participant always on the left of the screen and the opponent on the right. The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the opponent changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised opponents. We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a random time lapse (randomly chosen between 5 and 10 seconds).

```{r trustGameScreenshot,fig.cap = "Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee. The trustee in turn needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee makes a decision of how much to send back the investor", out.width="70%", out.height="70%", fig.align='center'}

@@

```{r grid,fig.cap = "Screenshot of the two axis grid in the Coaxing condition where participants were asked to report the valence and arousal of their emotional response ", out.width="70%", out.height="70%", fig.align='center'}

@@

On each round of the RTG, before deciding how much to return, participants were asked to provide feedback on the investment sent. This feedback consisted of a rating on two dimensions, provided via a two-dimensional field. What the axes of the field represented differed between the conditions. In the intervention condition, participants were asked to rate their emotional reaction to the other player’s choice of investment in terms of valence (positive or negative) and arousal (low or high). In the control condition, participants were asked to rate the choice of investment in terms of speed (slow or fast) and magnitude (low or high). Figure \ref{fig:grid} shows a screenshot of the feedback field in the intervention condition, where participants were asked about their emotional response to the investment. At the beginning of each game, participants were provided with detailed explanations of the meaning of the two axes as well as the opportunity to provide a baseline emotional state through using the field prior to the start of the game. Participants in the control condition were asked to chose any point on the two dimensional grid to show they have understood how to interact with the grid.

@@

# Self report questionnaires

@@

```{r postDefectionModel,fig.cap = "Summary of results of a Mixed-effects model fitted to participants returns restricted to  post-defection trials", fig.align='center'}

@@

```{r table-HMMs,fig.cap = "Table comparing the HMM-full models fit assuming the latent number of states is between 2 and 7. Using BIC, the 6 state model fits best", out.width="70%", out.height="70%", fig.align='center'}

@@

## Distribution of returns for each latent state for HMM model of participant’s returns

Focusing on the HMM-full models with 6 latent states (please see main manuscript for a definition) we estimated the investors' policy in each state assuming a discretised Gaussian response function as highlighted in the previous section (Figure \ref{fig:trusteeHMMPlot}). Each latent state was associated with a distribution over trustee actions. The latent states were ordered by the mean of the estimated Gaussian, such as higher numbered latent states reflect higher mean percentage return.

```{r trusteeHMMPlot,fig.cap = "Distribution of participants' percentage return for each of the latent states in the 6 state HMM-full model. The latent states are ordered by the mean of the Gaussian that best fits the policy in that state, so higher numbered states are more pro-social", fig.align='center'}

>>>>

Participants played a 15-round repeated trust game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. Each player was represented with an icon with the participant always on the left of the screen and the opponent on the right. The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the opponent changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised opponents. We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a random time lapse (randomly chosen between 5 and 10 seconds).

```{r trustGameScreenshot, include=F, fig.cap = "Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee. The trustee in turn needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee makes a decision of how much to send back the investor", out.width="70%", out.height="70%", fig.align='center'}

@@

```{r grid, include=F, fig.cap = "Screenshot of the two axis grid in the Coaxing condition where participants were asked to report the valence and arousal of their emotional response ", out.width="70%", out.height="70%", fig.align='center'}

@@

On each round of the RTG, before deciding how much to return, participants were asked to provide feedback on the investment sent. This feedback consisted of a rating on two dimensions, provided via a two-dimensional field. What the axes of the field represented differed between the conditions. In the intervention condition, participants were asked to rate their emotional reaction to the other player’s choice of investment in terms of valence (positive or negative) and arousal (low or high). In the control condition, participants were asked to rate the choice of investment in terms of speed (slow or fast) and magnitude (low or high). At the beginning of each game, participants were provided with detailed explanations of the meaning of the two axes as well as the opportunity to provide a baseline emotional state through using the field prior to the start of the game. Participants in the control condition were asked to chose any point on the two dimensional grid to show they have understood how to interact with the grid.

@@

# Self-report questionnaires

@@

```{r postDefectionModel,fig.cap = "Summary of results of a Mixed-effects model fitted to participants returns restricted to  post-defection trials", fig.align='center',fig.pos="H"}

@@

```{r table-HMMs,fig.cap = "Table comparing the HMM-full models fit assuming the latent number of states is between 2 and 7. Using BIC, the 6 state model fits best", out.width="70%", out.height="70%", fig.align='center',fig.pos="H"}

@@

```{r trusteeHMMPlot, include=F, fig.cap = "Distribution of participants' percentage return for each of the latent states in the 6 state HMM-full model. The latent states are ordered by the mean of the Gaussian that best fits the policy in that state, so higher numbered states are more pro-social", fig.align='center'}
